### 2024-03-02 examples.qmd
img uploads

---
#title: "Examples: A Brief Overview"
format:
  html:
    css: styles.css
    theme: cosmo
    #page-layout: full
    #grid:
      #sidebar-width: 250px
      #body-width: 2000px
      #margin-width: 0px
    #gutter-width: 1.5rem

---

## <i class="bi bi-book"></i> Examples

::: {.grid}
::: {.g-col-8}
::: {style="text-align: justify; font-size: 1.2em; line-stretch: 1.0; font-family: Georgia, serif;"}
[**Each Example**]{style="color: goldenrod; font-size: 1.2em; font-family: Verdana, sans-serif;"} uses data from the ABCD Study Dataset and provides a practical solution to a specific longitudinal method or analytic problem.

:::
:::

::: {.g-col-4}
```{=html}
<link rel="stylesheet" href="https://maxcdn.bootstrapcdn.com/bootstrap/4.3.1/css/bootstrap.min.css">

<div class="section_our_solution">
  <div class="row">
    <div class="col-lg-12 col-md-12 col-sm-12">
      <div class="our_solution_category">
        <div class="solution_cards_box">
          <div class="solution_card">
            <div class="hover_color_bubble"></div>
            <div class="so_top_icon">
              <svg id="Layer_1" enable-background="new 0 0 512 512" height="50" viewBox="0 0 512 512" width="40" xmlns="http://www.w3.org/2000/svg">
                <g>
                  <g>
                    <g>
                      <g>
                        <path d="m47.478 452.317 295.441 19.76c5.511.369 10.277-3.8 10.645-9.31l28.393-424.517c.369-5.511-3.8-10.276-9.31-10.645l-295.441-19.76c-5.511-.369-10.276 3.8-10.645 9.31l-28.394 424.517c-.368 5.511 3.8 10.277 9.311 10.645z" fill="#fae19e" />
                      </g>
                      <g>
                        <g>
                          <g>
                            <g>
                              <g>
                                <path d="m17.5 504.177h226.14l79.96-79.605v-355.86c0-5.523-4.477-10-10-10h-296.1c-5.523 0-10 4.477-10 10v425.466c0 5.522 4.477 9.999 10 9.999z" fill="#fff9e9" />
                              </g>
                              <path d="m313.601 58.712h-40c5.523 0 10 4.477 10 10v355.861l-.258 40.078 40.258-40.078v-355.861c0-5.523-4.477-10-10-10z" fill="#fff4d6" />
                            </g>
                          </g>
                        </g>
                        <path d="m243.64 504.177v-70.253c0-5.523 4.477-10 10-10h69.96z" fill="#ffeec2" />
                      </g>
                    </g>
                    <g>
                      <path d="m468.636 248.58-33.372.165v-50.826c0-9.183 7.463-16.662 16.673-16.708h.007c9.217-.046 16.693 7.371 16.693 16.562v50.807z" fill="#fed23a" />
                      <path d="m451.96 504.177c-10.362-10.277-16.196-24.263-16.208-38.857l-.062-73.973c0-.644.524-1.169 1.171-1.173l30.038-.149c.647-.003 1.171.517 1.171 1.161l.062 74.079c.012 14.531-5.749 28.472-16.015 38.756z" fill="#54b1ff" />
                      <path d="m451.959 469.333h-.01c-14.434.072-26.14-11.542-26.14-25.935v-213.527c0-6.778 5.477-12.283 12.255-12.316l27.626-.137c6.826-.034 12.378 5.49 12.378 12.316v213.436c0 14.38-11.687 26.091-26.109 26.163z" fill="#fdf385" />
                      <path d="m465.69 217.417-23.769.118c6.037.79 10.708 5.94 10.708 12.198v213.437c0 9.823-5.455 18.397-13.507 22.87 3.79 2.115 8.164 3.317 12.826 3.293h.01c14.422-.072 26.109-11.783 26.109-26.163v-213.436c.001-6.826-5.551-12.351-12.377-12.317z" fill="#faee6e" />
                      <path d="m491.274 247.925-71.615.355c-7.305.036-13.226 5.968-13.226 13.248 0 7.281 5.921 13.153 13.226 13.117l58.389-.29v77.489c0 7.281 5.921 13.153 13.226 13.117 7.305-.036 13.226-5.968 13.226-13.248v-90.672c0-7.28-5.922-13.152-13.226-13.116z" fill="#54b1ff" />
                      <g>
                        <path d="m491.274 247.925-38.441.188-.167 26.311 25.381-.067v77.489c0 7.281 5.921 13.153 13.226 13.117 7.305-.036 13.226-5.968 13.226-13.248v-90.672c.001-7.282-5.921-13.154-13.225-13.118z" fill="#3da7ff" />
                      </g>
                    </g>
                  </g>
                  <g fill="#060606">
                    <path d="m373.147 20.122-295.44-19.761c-9.631-.638-17.984 6.665-18.629 16.293l-2.311 34.557h-39.267c-9.649 0-17.5 7.851-17.5 17.5v425.466c0 9.649 7.851 17.5 17.5 17.5h226.141c1.96 0 3.902-.801 5.292-2.185l34.138-33.987c.347.074.701.133 1.065.157l58.282 3.898c9.302.614 18.005-6.952 18.629-16.293l28.393-424.515c.639-9.528-6.766-17.993-16.293-18.63zm-122.006 465.902v-52.1c0-1.378 1.122-2.5 2.5-2.5h51.9zm94.939-23.757c-.244 1.51-1.131 2.286-2.66 2.327l-46.28-3.096 31.752-31.611c1.414-1.407 2.209-3.32 2.209-5.315v-355.86c0-9.649-7.851-17.5-17.5-17.5h-77.993c-9.697 0-9.697 15 0 15h77.993c1.379 0 2.5 1.122 2.5 2.5v347.712h-62.46c-9.649 0-17.5 7.851-17.5 17.5v62.753h-218.641c-1.378 0-2.5-1.122-2.5-2.5v-425.465c0-1.378 1.122-2.5 2.5-2.5h178.168c9.697 0 9.697-15 0-15h-123.868l2.244-33.556c.244-1.511 1.131-2.286 2.661-2.327l295.44 19.76c1.511.244 2.287 1.131 2.328 2.661z" />
                    <path d="m267.827 237.047h-204.553c-4.142 0-7.5 3.358-7.5 7.5s3.358 7.5 7.5 7.5h204.553c4.143 0 7.5-3.358 7.5-7.5s-3.357-7.5-7.5-7.5z" />
                    <path d="m267.827 289.332h-204.553c-4.142 0-7.5 3.358-7.5 7.5s3.358 7.5 7.5 7.5h204.553c4.143 0 7.5-3.358 7.5-7.5s-3.357-7.5-7.5-7.5z" />
                    <path d="m55.774 192.262c0 4.142 3.358 7.5 7.5 7.5h204.553c4.143 0 7.5-3.358 7.5-7.5s-3.357-7.5-7.5-7.5h-204.553c-4.142 0-7.5 3.358-7.5 7.5z" />
                    <path d="m91.807 139.977c0 4.142 3.358 7.5 7.5 7.5h132.487c4.142 0 7.5-3.358 7.5-7.5s-3.358-7.5-7.5-7.5h-132.487c-4.142 0-7.5 3.358-7.5 7.5z" />
                    <path d="m194.755 438.787c-13.489.036-26.978.065-40.467.086-4.534.007-9.067.013-13.6.016-8.215.006-13.75-1.643-15.59-10.679-1.556-7.64-12.364-6.613-14.464 0-5.19 16.337-13.774 9.936-18.582-1.053-4.797-10.963-6.027-23.233-8.122-34.9-1.54-8.573-14.506-6.17-14.732 1.994-.298 10.751-1.302 21.331-4.031 31.758-2.815 10.758-7.034 21.097-11.222 31.376-3.651 8.961 10.867 12.816 14.464 3.988 3.711-9.108 7.427-18.266 10.193-27.714 5.14 12.36 15.774 26.34 30.927 18.101 2.819-1.533 5.452-3.712 7.763-6.253 7.88 9.106 19.609 8.388 30.584 8.375 15.627-.02 31.254-.054 46.881-.095 9.649-.025 9.667-15.025-.002-15z" />
                    <path d="m505.932 246.439c-3.897-3.878-9.255-5.867-14.695-6.014l-5.668.028v-10.719c0-6.529-3.878-13.427-9.433-16.862v-15.098c0-31.069-48.372-30.934-48.372.146v15.1c-5.659 3.498-9.455 9.741-9.455 16.852v10.982c-24.966 1.7-25.037 39.745.028 41.232.16 33.575.152 66.6-.028 100.737-.049 9.414 14.949 9.966 15 .079.18-34.166.188-67.22.029-100.823l37.211-.185s-.048 110.848-.048 160.784c0 24.338-37.219 24.5-37.219-.253l.013-13.677c.585-9.68-14.387-10.583-14.973-.904v12.834c0 11 3.402 20.316 9.988 26.869.586 15.693 7.198 30.878 18.369 41.956 3.205 3.18 7.642 2.208 10.744-.182 11.365-11.385 17.769-26.394 18.169-42.414 4.951-4.931 9.908-9.896 9.908-26.896l.006-68.351c12.97 3.689 26.494-6.348 26.494-19.946v-90.672c0-5.523-2.155-10.709-6.068-14.603zm-72.623-5.727v-10.841c0-2.219 1.523-4.08 3.573-4.633l30.025-.149c.84.208 1.615.605 2.243 1.231.915.911 1.419 2.123 1.419 3.414v10.794zm18.671-52c4.604 0 9.155 4.514 9.155 9.062v12.166l-18.372.091v-12.111c.001-5.053 4.133-9.183 9.217-9.208zm-.011 303.901c-3.487-4.942-6.009-10.531-7.417-16.406 2.322.503 4.674.765 7.027.765 2.627 0 5.253-.326 7.839-.957-1.374 5.964-3.892 11.587-7.449 16.598zm45.031-140.899c0 7.101-11.452 7.66-11.452.131 0 0 .013-70.974.021-77.48.005-4.196-3.483-7.509-7.558-7.509l-58.389.29c-7.242 0-7.073-11.331.074-11.366l71.615-.355c3.463.295 5.359 2.168 5.688 5.617v90.672z" />
                  </g>
                </g>
              </svg>
            </div>
            <div class="solu_description">
              <p>
                For an entry-level example of each longitudinal model
              </p>
              <button type="button" class="read_more_btn">Go Here</button>
            </div>
          </div>
          </div>
          </div>
        </div>
      </div>
    </div>
```
:::
:::

```{=html}
<section class="light">
	<div class="container py-2">
		<article class="postcard light blue">
			<a class="postcard__img_link" href="#">
				<img class="postcard__img" src="https://picsum.photos/1000/1000" alt="Image Title" />
			</a>
			<div class="postcard__text t-dark">
				<h1 class="postcard__title blue"><a href="#">Traditional Linear Models</a></h1>
				<div class="postcard__bar"></div>
				<ul class="postcard__tagbox">
					<li class="tag__item play blue">
						<a href="./1a_Examples_DifferenceScores_PairedTtests.qmd"><i class="fas fa-play
						mr-2"></i>Difference Scores (Paired T-test)</a>
						</li>
					<li class="tag__item play blue">
						<a href="./1b_Examples_DifferenceScores_SimpleRegression.qmd"><i class="fas fa-play
						mr-2"></i>Difference Scores (Simple Regression)</a>
						</li>
					<li class="tag__item play blue">
						<a href="./2_Examples_ResidualizedChangeScores.qmd"><i class="fas fa-play
						mr-2"></i>Residualized Change Scores</a>
						</li>
					<li class="tag__item play blue">
						<a href="./3a_Examples_LinearMixedModels.qmd"><i class="fas fa-play mr-2"></i>Linear Mixed
						Models (Random Intercept)</a>
					</li>
					<li class="tag__item play blue">
						<a href="./3b_Examples_LinearMixedModels.qmd"><i class="fas fa-play mr-2"></i>Linear Mixed
						Models (Random Slope)</a>
					</li>
				</ul>
			</div>
		</article>
		<article class="postcard light red">
			<a class="postcard__img_link" href="#">
				<img class="postcard__img" src="https://picsum.photos/501/500" alt="Image Title" />	
			</a>
			<div class="postcard__text t-dark">
				<h1 class="postcard__title red"><a href="#">Traditional Nonlinear Models</a></h1>
				<div class="postcard__bar"></div>
				<div class="postcard__preview-txt">
				<ul class="postcard__tagbox">
				  <li class="tag__item play blue">
						<a href="./_4_Examples_SignedRankTest.qmd"><i class="fas fa-play mr-2"></i>Signed-Rank
						Test</a>
						</li>
					<li class="tag__item play blue">
						<a href="./_5_Examples_MarginalModels.qmd"><i class="fas fa-play mr-2"></i>Marginal Models</a>
						</li>
						<li class="tag__item play blue">
						<a href="./_6_Examples_GeneralizedEstimatingEquations.qmd"><i class="fas fa-play
						mr-2"></i>Generalized
						Estimating Equations</a>
						</li>
						<li class="tag__item play blue">
						<a href="./_7_Examples_GeneralizedLinearMixedEffectsModels.qmd"><i class="fas fa-play
						mr-2"></i>Generalized Linear
						Mixed Effects Models</a>
						</li>
				</ul>
			</div>
		</article>
		<article class="postcard light red">
			<a class="postcard__img_link" href="#">
				<img class="postcard__img" src="https://picsum.photos/601/600" alt="Image Title" />	
			</a>
			<div class="postcard__text t-dark">
				<h1 class="postcard__title red"><a href="#">Structural Equation Models</a></h1>
				<div class="postcard__bar"></div>
				<div class="postcard__preview-txt">
				<ul class="postcard__tagbox">
				  <li class="tag__item play blue">
						<a href="./_9_Examples_LatentChangeScoresModels.qmd"><i class="fas fa-play
						mr-2"></i>Latent Change Score Models</a>
						</li>
					<li class="tag__item play blue">
						<a href="./_10_Examples_LatentGrowthCurveModels.qmd"><i class="fas fa-play
						mr-2"></i>Latent Growth Curve Models</a>
						</li>
					<li class="tag__item play blue">
						<a href="./_12_Examples_LatentTransitionAnalysis.qmd"><i class="fas fa-play
						mr-2"></i>Latent Transition Analysis</a>
						</li>
				</ul>
			</div>
		</article>
		<article class="postcard light red">
			<a class="postcard__img_link" href="#">
				<img class="postcard__img" src="https://picsum.photos/500/653" alt="Image Title" />	
			</a>
			<div class="postcard__text t-dark">
				<h1 class="postcard__title red"><a href="#">Advanced SEM</a></h1>
				<div class="postcard__bar"></div>
				<div class="postcard__preview-txt">
				<ul class="postcard__tagbox">
				  <li class="tag__item play blue">
						<a href="./_15_Examples_RandomInterceptCrosslaggedPanelModels.qmd"><i class="fas fa-play
						mr-2"></i>Random-Intercept Crosslagged Panel Model</a>
						</li>
						<li class="tag__item play blue">
						<a href="./_16_Examples_LatentCurveModelsStructuredResiduals.qmd"><i class="fas fa-play
						mr-2"></i>Latent Curve Models with Structured Residuals</a>
						</li>
				</ul>
			</div>
		</article>
		<article class="postcard light red">
			<a class="postcard__img_link" href="#">
				<img class="postcard__img" src="https://picsum.photos/801/800" alt="Image Title" />	
			</a>
			<div class="postcard__text t-dark">
				<h1 class="postcard__title red"><a href="#">Additional Models</a></h1>
				<div class="postcard__bar"></div>
				<div class="postcard__preview-txt">
				<ul class="postcard__tagbox">
				  <li class="tag__item"><i class="fas fa-tag mr-2"></i></li>
					<li class="tag__item"><i class="fas fa-clock mr-2"></i></li>
				</ul>
			</div>
		</article>

	</div>
</section>
```

```{=html}
<div class="d-flex flex-column justify-content-center w-100 h-100">

	<div class="d-flex flex-column justify-content-center align-items-center">
		<h1 class="fw-light text-white m-0">Pure CSS Gradient Background Animation</h1>
		<div class="btn-group my-5">
			<a href="https://codepen-api-export-production.s3.us-west-2.amazonaws.com/zip/PEN/pyBNzX/1578778289271/pure-css-gradient-background-animation.zip" class="btn btn-outline-light" aria-current="page"><i class="fas fa-file-download me-2"></i> SOURCE CODE</a>
			<a href="https://codepen.io/P1N2O/full/pyBNzX" class="btn btn-outline-light">FULL SCREEN <i class="fas fa-expand ms-2"></i></a>
		</div>
		<a href="https://manuel.pinto.dev" class="text-decoration-none">
			<h5 class="fw-light text-white m-0">— Pen by Manuel Pinto —</h5>
		</a>
	</div>
</div>
</div>
```