### 2024-03-02 posts/_10_Examples_LatentGrowthCurveModels.qmd

---
title: "Latent Growth Curve Models"
author: "Biostatistics Working Group"
---

# XXXXXXXXXXX {.unnumbered}

## Overview

<p>

XXXXXXXXXXX Models xxxxxxx. This xxxxxxx.

In this example, we will utilize xxxxxx models to analyze xxxxxx obtained across multiple measurement occasions for a sample of youth participating in the ABCD Study. Our primary objective is to understand xxxxxx, while factoring in the clustered nature of observations within individuals over time. XXXX models facilitate this by xxxxx.

</p>

## Preliminary Setup

::: panel-tabset
### Install Packages {.tabset .tabset-fade .tabset-pills}

::: blue
> **This code installs the r packages necessary for this example, if
> they are not already installed**

```{r pckg-install}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

# Create a list of required packages
packages_required <- c("tidyverse","rstatix","DT")

# Check which packages are not installed and install them
packages_to_install <- setdiff(packages_required, rownames(installed.packages()))
if (length(packages_to_install) > 0) {
    install.packages(packages_to_install)
}

# Load the required packages
lapply(packages_required, library, character.only = TRUE)

```
:::

### Load Packages

::: blue
**This code loads the r libraries necessary for this example**

```{r lib-load}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

library(tidyverse)    # Collection of R packages for data science
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(DT)           # Rendering interactive data tables

```
:::

### Config Options

::: blue
**This code configures knitr code chunk options**

```{r config}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, 
                      comment=NA, cache=T, code_folding=T,
                      R.options=list(width=220, digits = 3),
                      fig.align='center', 
                      out.width='75%', fig.asp=.75)
```
:::
:::

## Descriptives Overview

::: panel-tabset
### Read and View Data {.tabset .tabset-fade .tabset-pills}

::: blue
**This code reads in and shows the data to be used in the current
example**

```{r read-data}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE
#| cache: FALSE

# Set the data paths
data_path_1 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/demo5.0.rds"
data_path_2 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/core-rds-5.0/non-imaging_excluding_nt_5.0.rds"

# Read the data
data_demographics <- readRDS(data_path_1)
data_nonimaging <- readRDS(data_path_2)

# Subset the nonimaging data to include desired variables
selected_vars <- c("src_subject_id", "eventname", "nihtbx_totalcomp_fc", "anthroweightcalc", "anthroheightcalc")
subset_data <- data_nonimaging[, selected_vars]

library(dplyr)
# # Merge the datasets on 'src_subject_id' and 'eventname'
merged_data <- data_demographics %>%
  full_join(subset_data, by = c("src_subject_id", "eventname"))

# Inspect the merged data structure
str(merged_data)

# Define event names to be retained in the analysis and convert variables to appropriate data types
eventnames_to_include <- c("baseline_year_1_arm_1",
                           "1_year_follow_up_y_arm_1",
                           "2_year_follow_up_y_arm_1",
                           "3_year_follow_up_y_arm_1",
                           "4_year_follow_up_y_arm_1")

df <- merged_data %>%
  filter(eventname %in% eventnames_to_include) %>%
  mutate(
    src_subject_id = as.factor(src_subject_id),
    eventname = factor(eventname, levels = eventnames_to_include, ordered = TRUE),
    age = as.numeric(age),
    sex = as.factor(sex),
    race.4level = as.factor(race.4level),
    hisp = as.factor(hisp),
    high.educ.bl = as.factor(high.educ.bl),
    household.income.bl = as.factor(household.income.bl),
    acs_raked_propensity_score = as.numeric(acs_raked_propensity_score),
    rel_family_id.bl = as.factor(rel_family_id.bl),
    site_id_l = as.factor(site_id_l),
    nihtbx_totalcomp_fc = as.numeric(nihtbx_totalcomp_fc),
    anthroweightcalc = as.numeric(anthroweightcalc),
    anthroheightcalc = as.numeric(anthroheightcalc)
  ) %>%
  # Exclude cases from unused assessment waves
  filter(!is.na(eventname))

```
:::

### Descriptives

::: blue
**This code creates a descriptives table**

```{r descriptives}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

# Define a function to compute descriptives
compute_descriptives <- function(data, event_name) {
  # For factor variables
  sex_desc <- paste0(table(data$sex), " (", round(100 * prop.table(table(data$sex)), 1), "%)")
  race_desc <- paste0(table(data$race.4level), " (", round(100 * prop.table(table(data$race.4level)), 1), "%)")

  # For numeric variables
  age_desc <- paste0(round(mean(data$age, na.rm = TRUE), 2), " (", round(sd(data$age, na.rm = TRUE), 2), ")")
  weight_desc <- paste0(round(mean(data$anthroweightcalc, na.rm = TRUE), 2), " (", round(sd(data$anthroweightcalc, na.rm = TRUE), 2), ")")
  height_desc <- paste0(round(mean(data$anthroheightcalc, na.rm = TRUE), 2), " (", round(sd(data$anthroheightcalc, na.rm = TRUE), 2), ")")

  # Combine into a data frame
  desc_df <- data.frame(
    Variable = c("Sex - Female", "Sex - Male or other", "Race - Asian", "Race - Black", "Race - Other/Mixed", "Race - White", "Age", "Weight", "Height"),
    Value = c(sex_desc, race_desc, age_desc, weight_desc, height_desc)
  )

  # Rename the Value column based on event name
  colnames(desc_df)[2] <- event_name
  return(desc_df)
}

# Compute descriptives for each event
baseline_desc <- compute_descriptives(subset(df, eventname == "baseline_year_1_arm_1"), "baseline_year_1_arm_1")
one_year_desc <- compute_descriptives(subset(df, eventname == "1_year_follow_up_y_arm_1"), "1_year_follow_up_y_arm_1")
two_year_desc <- compute_descriptives(subset(df, eventname == "2_year_follow_up_y_arm_1"), "2_year_follow_up_y_arm_1")
three_year_desc <- compute_descriptives(subset(df, eventname == "3_year_follow_up_y_arm_1"), "3_year_follow_up_y_arm_1")

# Join all data frames side-by-side
final_table <- baseline_desc %>%
  left_join(one_year_desc, by = "Variable") %>%
  left_join(two_year_desc, by = "Variable") %>%
  left_join(three_year_desc, by = "Variable")

# Adjust for the required format
final_table[1:6, 3:5] <- NA

# Round numeric values to two decimal places
numeric_cols <- sapply(final_table, is.numeric)
final_table[numeric_cols] <- lapply(final_table[numeric_cols], round, 2)

# Create heading rows with the same columns as final_table
heading_rows <- data.frame(
  Variable = c("Sex", "Race"),
  baseline_year_1_arm_1 = NA_real_,
  `1_year_follow_up_y_arm_1` = NA_real_,
  `2_year_follow_up_y_arm_1` = NA_real_,
  `3_year_follow_up_y_arm_1` = NA_real_
)

# Set column names of heading_rows to match final_table
colnames(heading_rows) <- colnames(final_table)

# Introduce group labels and adjust the "Variable" column
final_table <- rbind(
                    heading_rows[1,], 
                    final_table[1:2,],
                    heading_rows[2,], 
                    final_table[3:6,], 
                    final_table[7:9,]
                    )

# Update the Variable column to remove redundant factor variable name
final_table$Variable <- gsub("Sex - ", "", final_table$Variable)
final_table$Variable <- gsub("Race - ", "", final_table$Variable)
final_table$Variable[final_table$Variable == "Male or other"] <- "Male"

# Add non-breaking spaces for increased indentation
final_table$Variable[final_table$Variable %in% c("Female", "Male", "Asian", "Black", "Other/Mixed", "White")] <- 
  paste0(rep("&nbsp;", 6), final_table$Variable[final_table$Variable %in% c("Female", "Male", "Asian", "Black", "Other/Mixed", "White")])

# Update column headers
colnames(final_table)[2:5] <- c("Baseline", "Year 1", "Year 2", "Year 3")

# Display the table interactively without row names, with updated column headers, and with HTML entities rendered
datatable(final_table, 
          colnames = c("", "Baseline", "Year 1", "Year 2", "Year 3"),
          options = list(pageLength = nrow(final_table), autoWidth = TRUE),
          rownames = FALSE, escape = FALSE,
          caption = "Descriptives Table")  # Add table title

```

:::
:::

## Results 
<div class = "blue">

::: panel-tabset
### Build Model {.tabset .tabset-fade .tabset-pills}

::: blue

------

**STEP 1: Convert dataframe from long-to-wide format**
```{r long-to-wide}
#| class.source: 'fold-hide'
#| message: FALSE
#| warning: FALSE
#| echo: TRUE

# Split the data into static (unchanging) and changing variables
# Static variables are taken only from the baseline measurement occasion
df_static <- df %>%
  filter(eventname == "baseline_year_1_arm_1") %>%
  select(src_subject_id, site_id_l, sex)

df_changing <- df %>%
  select(src_subject_id, eventname, anthroweightcalc, anthroheightcalc)

# Convert the changing variables from long format to wide format
df_wide_changing <- df_changing %>%
  gather(variable, value, -src_subject_id, -eventname) %>%
  unite("variable_event", variable, eventname) %>%
  spread(variable_event, value)

# Merge the static and changing data frames based on the 'src_subject_id' variable
df_wide <- df_static %>%
  left_join(df_wide_changing, by = "src_subject_id")

# Rename columns to shorter names for easier reference
df_wide <- df_wide %>% 
        rename("anthroweightcalc_0"="anthroweightcalc_baseline_year_1_arm_1",
               "anthroweightcalc_1"="anthroweightcalc_1_year_follow_up_y_arm_1",
               "anthroweightcalc_2"="anthroweightcalc_2_year_follow_up_y_arm_1",
               "anthroweightcalc_3"="anthroweightcalc_3_year_follow_up_y_arm_1",
               "anthroweightcalc_4"="anthroweightcalc_4_year_follow_up_y_arm_1",
               "anthroheightcalc_0"="anthroheightcalc_baseline_year_1_arm_1",
               "anthroheightcalc_1"="anthroheightcalc_1_year_follow_up_y_arm_1",
               "anthroheightcalc_2"="anthroheightcalc_2_year_follow_up_y_arm_1",
               "anthroheightcalc_3"="anthroheightcalc_3_year_follow_up_y_arm_1",
               "anthroheightcalc_4"="anthroheightcalc_4_year_follow_up_y_arm_1"
               )

# Reorder columns in the dataframe
df_wide <- df_wide %>% select(src_subject_id, site_id_l, sex, anthroweightcalc_0, anthroweightcalc_1, anthroweightcalc_2, anthroweightcalc_3, anthroweightcalc_4, anthroheightcalc_0, anthroheightcalc_1, anthroheightcalc_2, anthroheightcalc_3, anthroheightcalc_4)

```

The code fits a xxxx to examine the 'xxxxx' variable across time points ('eventname'). It incorporates xxxxxx. The results of the model are then printed to provide a detailed summary of the fitted model parameters.

**STEP 2: Compute xxxxx**

```{r}

# Define the Latent Growth Curve Model
model <- '
  # Latent variables
    i =~ 1 * anthroheightcalc_0 + 1 * anthroheightcalc_1 + 1 * anthroheightcalc_2
    s =~ 0 * anthroheightcalc_0 + 1 * anthroheightcalc_1 + 2 * anthroheightcalc_2

  # Estimate means of latent variables
    i ~ 1
    s ~ 1

  # Variances
    i ~~ i
    s ~~ s

  # Covariance
    i ~~ s
'
```

```{r}
# Fit the model
fit <- sem(model, data = df_wide, missing = "FIML")

summary(fit, fit.measures=TRUE, standardized=TRUE, rsquare=TRUE)

```