### 2024-03-02 posts/_4_Examples_SignedRankTest.qmd

---
title: "Signed-Rank Test"
author: "Biostatistics Working Group"
---

# Signed-Rank Test {.unnumbered}

####
It is actually quite common that numeric depend variables need to be transformed or converted into ranks, i.e. ordinal variables, because the distribution of residuals does not allow the application of parametric tests such as t-tests or linear regression. In such cases, as we are dealing with rank (ordinal) data, the application of a chi-square test is unwarranted and we need to use another test. There are different alternatives depending on whether the data are paired (coming from the same individuals) or if all observations are independent.

The non-parametric alternative for independent t-tests, i.e. for data where we are dealing with two separate groups and a numeric dependent variable that violates parametric assumptions (or an ordinal dependent variable), is the Mann-Whitney U-test. In contrast, if the groups under investigation represent identical participants that are tested under two conditions, the appropriate alternative is a Wilcoxon Signed Rank test (which is thus the alternative for paired t-test).

Imagine we wanted to determine if two language families differed with respect to the size of their phoneme inventories. You have already ranked the inventory sizes and would now like to now if language family correlates with inventory size. As such, we are dealing with two independent groups and we want to implement a non-parametric alternative of a t-test. To answer this question, you create the table shown below.
####

## Overview

<p>

The Signed-Rank Test xxxxxxx. This xxxxxxx.

In this example, we will utilize the signed-rank test to analyze xxxxxx obtained across multiple measurement occasions for a sample of youth participating in the ABCD Study. Our primary objective is to understand xxxxxx, while factoring in the clustered nature of observations within individuals over time. The signed-rank test facilitates this by xxxxx.

</p>

## Preliminary Setup

::: panel-tabset
### Install Packages {.tabset .tabset-fade .tabset-pills}

::: blue
> **This code installs the r packages necessary for this example, if
> they are not already installed**

```{r pckg-install}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

# Create a list of required packages
packages_required <- c("tidyverse","rstatix","DT", "gtsummary","easystats")

# Check which packages are not installed and install them
packages_to_install <- setdiff(packages_required, rownames(installed.packages()))
if (length(packages_to_install) > 0) {
    install.packages(packages_to_install)
}

# Load the required packages
lapply(packages_required, library, character.only = TRUE)

```
:::

### Load Packages

::: blue
**This code loads the r libraries necessary for this example**

```{r lib-load}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

library(tidyverse)    # Collection of R packages for data science
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(DT)           # Rendering interactive data tables
library(easystats)    #
library(gtsummary)    #
```
:::

### Config Options

::: blue
**This code configures knitr code chunk options**

```{r config}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, 
                      comment=NA, cache=T, code_folding=T,
                      R.options=list(width=220, digits = 3),
                      fig.align='center', 
                      out.width='75%', fig.asp=.75)
```
:::
:::

## Descriptives Overview

::: panel-tabset
### Read and View Data {.tabset .tabset-fade .tabset-pills}

::: blue
**This code reads in and shows the data to be used in the current
example**

```{r read-data}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE
#| cache: FALSE

# Set the data paths
data_path_1 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/demo5.0.rds"
data_path_2 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/core-rds-5.0/non-imaging_excluding_nt_5.0.rds"

# Read the data
data_demographics <- readRDS(data_path_1)
data_nonimaging <- readRDS(data_path_2)

# Subset the nonimaging data to include desired variables
selected_vars <- c("src_subject_id", "eventname", "anthroweightcalc")
subset_data <- data_nonimaging[, selected_vars]

library(dplyr)
# # Merge the datasets on 'src_subject_id' and 'eventname'
merged_data <- data_demographics %>%
  full_join(subset_data, by = c("src_subject_id", "eventname"))

# Inspect the merged data structure
str(merged_data)

# Define event names to be retained in the analysis and convert variables to appropriate data types
eventnames_to_include <- c("baseline_year_1_arm_1",
                           "1_year_follow_up_y_arm_1",
                           "2_year_follow_up_y_arm_1",
                           "3_year_follow_up_y_arm_1",
                           "4_year_follow_up_y_arm_1")

df <- merged_data %>%
  filter(eventname %in% eventnames_to_include) %>%
  mutate(
    src_subject_id = as.factor(src_subject_id),
    eventname = factor(eventname, levels = eventnames_to_include, ordered = TRUE),
    age = as.numeric(age),
    sex = as.factor(sex),
    race.4level = as.factor(race.4level),
    hisp = as.factor(hisp),
    high.educ.bl = as.factor(high.educ.bl),
    household.income.bl = as.factor(household.income.bl),
    acs_raked_propensity_score = as.numeric(acs_raked_propensity_score),
    rel_family_id.bl = as.factor(rel_family_id.bl),
    site_id_l = as.factor(site_id_l),
    anthroweightcalc = as.numeric(anthroweightcalc)
  ) %>%
  # Exclude cases from unused assessment waves
  filter(!is.na(eventname))

#####
#Creating new ordinal variable for the example
df$education <- factor(
  with(df, ifelse(high.educ.bl %in% c("1st grade", "2nd grade", "3rd grade", "4th grade",
                                      "5th grade",
                                      "6th grade", "7th grade", "8th grade", "9th grade",
                                      "10th grade",
                                      "11th grade", "12th grade"), "lt_highschool",
                  ifelse(high.educ.bl == "High school graduate", "highschool",
                  ifelse(high.educ.bl == "GED or equivalent Diploma", "GED",
                  ifelse(high.educ.bl == "Some college", "some_college",
                  ifelse(high.educ.bl %in% c("Associate degree: Academic Program", 
                                              "Associate degree: Occupational"),
                         "associate_degree",
                  ifelse(high.educ.bl == "Bachelor's degree (ex. BA", "Bachelor's_degree",
                  ifelse(high.educ.bl == "Master's degree (ex. MA", "Master's_degree",
                  ifelse(high.educ.bl %in% c("Doctoral degree (ex. PhD", "Professional
                                             School degree (ex. MD"), 
                         "Advanced_degree", NA))))))))))
  levels = c("lt_highschool", "highschool", "GED", "some_college", "associate_degree",
             "Bachelor's_degree", "Master's_degree", "Advanced_degree")
  ordered = TRUE

####

####
df$income <- factor(
  with(df, ifelse(household.income.bl == "Less than $5,000", "<$5,000",
                  ifelse(household.income.bl == "$5,000 through $11,999", "$5,000-$11,999",
                  ifelse(household.income.bl == "$12,000 through $15,999", "$12,000-$15,999",
                  ifelse(household.income.bl == "$16,000 through $24,999", "$16,000-$24,999",
                  ifelse(household.income.bl == "$25,000 through $34,999", "$25,000-$34,999",
                  ifelse(household.income.bl == "$35,000 through $49,999", "$35,000-$49,999",
                  ifelse(household.income.bl == "$50,000 through $74,999", "$50,000-$74,999",
                  ifelse(household.income.bl == "$75,000 through $99,999", "$75,000-$99,999",
                  ifelse(household.income.bl == "$100,000 through $199,999", "$100,000-$199,999",
                  ifelse(household.income.bl == "$200,000 and greater", "≥$200,000",
                  NA))))))))))),
  levels = c("<$5,000", "$5,000-$11,999", "$12,000-$15,999", "$16,000-$24,999",
             "$25,000-$34,999", "$35,000-$49,999", "$50,000-$74,999",
             "$75,000-$99,999", "$100,000-$199,999", "≥$200,000"),
  ordered = TRUE
)

####

# Create a mapping from factor levels to numeric values
income_levels <- c("Less than $5,000", "$5,000 through $11,999", "$12,000 through $15,999", "$16,000 through $24,999", "$25,000 through $34,999", "$35,000 through $49,999", "$50,000 through $74,999", "$75,000 through $99,999", "$100,000 through $199,999", "$200,000 and greater", "Don't know", "Refuse to answer", "No deseo responder")

# We assign numbers 1 through 10, but we have 12 levels, including "Don't know", "Refuse to answer", and "No deseo responder"
# Let's assume "Don't know" and any refusal to answer will be treated as NA
income_numeric_values <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, NA, NA, NA)

# Create a named vector for easy lookup
income_values <- setNames(income_numeric_values, income_levels)

# Convert the `income` factor to numeric using the named vector
df$income_numeric <- income_values[as.character(df$household.income.bl)]

# Handling NAs and refused to answer
df$income_numeric[is.na(df$household.income.bl) | df$household.income.bl %in% c("Don't know", "Refuse to answer", "No deseo responder")] <- NA

```
:::

### Descriptives

::: blue
**This code creates a descriptives table**

```{r descriptives}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

descriptives_table <- df %>%
  select(eventname, sex, race.4level, income_numeric, education) %>%
  mutate(eventname = factor(eventname, labels = c("Baseline", "Year 1","Year 2","Year 3","Year 4"))) %>%
  mutate(sex = factor(sex, labels = c("Female", "Male"))) %>%
  tbl_summary(
    by = eventname,
    missing = "no",
    label = list(sex ~ "Sex", race.4level ~ "Race", income_numeric ~ "Income", 
                 education ~ "Education"),
    statistic = list(all_continuous() ~ "{mean} ({sd}) )", all_categorical() ~ "{p}%"),
  ) %>%
modify_header(all_stat_cols() ~ "**{level}**<br>N = {n}") %>%
  bold_labels() %>%
  italicize_levels() %>%
  modify_spanning_header(all_stat_cols() ~ "**Assessment Wave**")
theme_gtsummary_compact()

descriptives_table
```

:::
:::

## Results 
<div class = "blue">

::: panel-tabset
### Build Model {.tabset .tabset-fade .tabset-pills}

::: blue

The code fits a xxxx to examine the 'xxxxx' variable across time points ('eventname'). It incorporates xxxxxx. The results of the model are then printed to provide a detailed summary of the fitted model parameters.

**STEP 1: Convert dataframe from long-to-wide format**
```{r long-to-wide}
#| class.source: 'fold-hide'
#| message: FALSE
#| warning: FALSE
#| echo: TRUE

# Split the data into static (unchanging) and changing variables
# Static variables are taken only from the baseline measurement occasion
df_static <- df %>%
  filter(eventname == "baseline_year_1_arm_1") %>%
  select(src_subject_id, site_id_l, sex)

df_changing <- df %>%
  select(src_subject_id, eventname, income_numeric, education)

# Convert the changing variables from long format to wide format
df_wide_changing <- df_changing %>%
  gather(variable, value, -src_subject_id, -eventname) %>%
  unite("variable_event", variable, eventname) %>%
  spread(variable_event, value)

# Merge the static and changing data frames based on the 'src_subject_id' variable
df_wide <- df_static %>%
  left_join(df_wide_changing, by = "src_subject_id")

# Rename columns to shorter names for easier reference
df_wide <- df_wide %>% 
        rename("income_numeric_0"="income_numeric_baseline_year_1_arm_1",
               "income_numeric_1"="income_numeric_1_year_follow_up_y_arm_1",
               "income_numeric_2"="income_numeric_2_year_follow_up_y_arm_1",
               "income_numeric_3"="income_numeric_3_year_follow_up_y_arm_1",
               "income_numeric_4"="income_numeric_4_year_follow_up_y_arm_1"
               )

# Reorder columns in the dataframe
df_wide <- df_wide %>% select(src_subject_id, site_id_l, sex, income_numeric_0, income_numeric_1, income_numeric_2, income_numeric_3, income_numeric_4)

```

**Descriptive Statistics for the Income Variable**
```{r}
## Summary statistics
# Compute summary statistics for Income by eventname
summary <- df %>%
    group_by(eventname) %>%
    get_summary_stats(income_numeric, type = "mean_sd")

data.frame(summary)
```

The summary statistics provide insights into the Weight variable across different `eventname` categories. At baseline, the average weight is approximately `r summary$mean[summary$eventname == "baseline_year_1_arm_1"]` units with a standard deviation of `r summary$sd[summary$eventname == "baseline_year_1_arm_1"]`. Over the years, there’s a noticeable increase in average weight: by the first year, it’s about `r summary$mean[summary$eventname == "1_year_follow_up_y_arm_1"]` units, and it continues to rise, reaching approximately `r summary$mean[summary$eventname == "4_year_follow_up_y_arm_1"]` units by the fourth year. The standard deviation remains relatively consistent over the years, suggesting similar variability in weight measurements across these time points.

**STEP 2: Compute Signed Rank Test**
```{r}

#########
df_wide$income_numeric_0 <- as.numeric(as.character(df_wide$income_numeric_0))
df_wide$income_numeric_1 <- as.numeric(as.character(df_wide$income_numeric_1))
df_wide$income_numeric_4 <- as.numeric(as.character(df_wide$income_numeric_4))
df_wide_complete <- na.omit(df_wide[, c("income_numeric_0", "income_numeric_4")])

model <- wilcox.test(df_wide_complete$income_numeric_0, df_wide_complete$income_numeric_4, paired = TRUE)

report(wilcox.test(df_wide_complete$income_numeric_0, df_wide_complete$income_numeric_4))

```

**Descriptive Statistics for the Signed Rank Test result**
```{r model-output}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| code-summary: testing

library(gridExtra)

# Creating the first density plot with ggplot2
p1 <- ggplot(data.frame(x = na.omit(df_wide$income_numeric_0)), aes(x = x)) +
  geom_density(fill="red", color="blue") +
  labs(title = "Kernel Density of Income at Baseline")

# Creating the second density plot with ggplot2
p2 <- ggplot(data.frame(x = na.omit(df_wide$income_numeric_1)), aes(x = x)) +
  geom_density(fill="blue", color="red") +
  labs(title = "Kernel Density of Income at T1")

p3<- ggplot(df_wide, aes(x = income_numeric_0, y = income_numeric_1)) + geom_point(color="#04c24a")

# Create both plot and store in variable
p4<-ggplot(df_wide, aes(x=income_numeric_0, y=income_numeric_1))+
geom_segment( aes(x=income_numeric_0, xend=income_numeric_1, y=0, yend=income_numeric_2), color="grey") + geom_point( color="green", size=4)

p5<-ggplot(df_wide, aes(x = income_numeric_0)) + 
geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8)

p6<-ggplot(df_wide, aes(x=income_numeric_0, y=income_numeric_1)) + 
geom_bar(fill="#69b3a2", stat = "identity")

# Arrange the ggplot2 objects on a grid
grid.arrange(p1, p2, p3, p4, p5, p6, ncol = 3)

ggplot(df_wide, aes(x = sex, y = income_numeric_0, fill = sex)) +
  geom_boxplot() +
  scale_fill_manual(values = c("orange", "darkgrey")) +
  theme_bw() +
  theme(legend.position = "none")
```

This signed rank test evaluates xxxx. The model output provides:

i. xxxxx;
ii. xxxxx;
iii. xxxxx;
iv. xxxxx;
v. xxxxx.

With a xxxx of `xxxxx` and a p-value of `xxxxx`, we can xxxx. Furthermore, we can xxx that xxxxx. This indicates xxxxxx in weight from Baseline to Year 1.