### 2024-03-04 index.qmd
update index

---
#title: "Longitudinal Analysis"

format:
  html:
    css: styles.css
    #theme: cosmo
    page-layout: custom
    grid:
      sidebar-width: 200px
      body-width: 2000px
      margin-width: 0px
      gutter-width: 0rem
---

Test