### 2024-03-06 _posts/2_Examples_ResidualizedChangeScores.qmd
revised manuscript docs

---
title: "Residualized Change Scores"
author: "Biostatistics Working Group"
---

# Mulitple Regression {.unnumbered}

## Overview

<p>

This example assesses whether height, measured at the 1-Year follow-up (T1) in a subsample of ABCD participants, is predicted by weight and height measured at baseline (T0). A visual inspection is further conducted via a scatterplot to graphically represent the relationship between height at Year_1 and weight at baseline, controlling for height at baseline. The ensuing analysis and interpretations are detailed in the subsequent sections.

</p>

## Preliminary Setup

::: panel-tabset
### Install Packages {.tabset .tabset-fade .tabset-pills}

::: blue
> **This code installs the r packages necessary for this example, if
> they are not already installed**

```{r pckg-install}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

# Create a list of required packages
packages_required <- c("tidyverse","easystats","lme4","gtsummary","report","broom","gridExtra")

# Check which packages are not installed and install them
packages_to_install <- setdiff(packages_required, rownames(installed.packages()))
if (length(packages_to_install) > 0) {
    install.packages(packages_to_install)
}

# Load the required packages
lapply(packages_required, library, character.only = TRUE)

```
:::

### Load Packages

::: blue
**This code loads the r libraries necessary for this example**

```{r lib-load}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

library(tidyverse)    # Collection of R packages for data science
library(easystats)    # Collection of R packages statistical modeling, visualization, and reporting
library(gtsummary)    # Publication-ready tables
library(report)       # Easy reporting of regression analyses
library(broom)        # Tidy and augment statistical models output
library(gridExtra)    # Arrange multiple grid-based plots on a page

```
:::

### Config Options

::: blue
**This code configures knitr code chunk options**

```{r config}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, 
                      comment=NA, cache=T, code_folding=T,
                      R.options=list(width=220, digits = 3),
                      fig.align='center', 
                      out.width='75%', fig.asp=.75)
```
:::
:::

## Descriptives Overview

::: panel-tabset
### Read and View Data {.tabset .tabset-fade .tabset-pills}

::: blue
**This code reads in and shows the data to be used in the current
example**

```{r read-data}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE
#| cache: FALSE

# Set the data paths
data_path_1 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/demo5.0.rds"
data_path_2 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/core-rds-5.0/non-imaging_excluding_nt_5.0.rds"

# Read the data
data_demographics <- readRDS(data_path_1)
data_nonimaging <- readRDS(data_path_2)

# Subset the nonimaging data to include desired variables
selected_vars <- c("src_subject_id", "eventname", "nihtbx_totalcomp_fc", "anthroweightcalc", "anthroheightcalc")
subset_data <- data_nonimaging[, selected_vars]

library(dplyr)
# # Merge the datasets on 'src_subject_id' and 'eventname'
merged_data <- data_demographics %>%
  full_join(subset_data, by = c("src_subject_id", "eventname"))

# Inspect the merged data structure
str(merged_data)

# Define event names to be retained in the analysis and convert variables to appropriate data types
eventnames_to_include <- c("baseline_year_1_arm_1",
                           "1_year_follow_up_y_arm_1",
                           "2_year_follow_up_y_arm_1",
                           "3_year_follow_up_y_arm_1",
                           "4_year_follow_up_y_arm_1")

df <- merged_data %>%
  filter(eventname %in% eventnames_to_include) %>%
  mutate(
    src_subject_id = as.factor(src_subject_id),
    eventname = factor(eventname, levels = eventnames_to_include, ordered = TRUE),
    age = as.numeric(age),
    sex = as.factor(sex),
    race.4level = as.factor(race.4level),
    hisp = as.factor(hisp),
    high.educ.bl = as.factor(high.educ.bl),
    household.income.bl = as.factor(household.income.bl),
    acs_raked_propensity_score = as.numeric(acs_raked_propensity_score),
    rel_family_id.bl = as.factor(rel_family_id.bl),
    site_id_l = as.factor(site_id_l),
    nihtbx_totalcomp_fc = as.numeric(nihtbx_totalcomp_fc),
    anthroweightcalc = as.numeric(anthroweightcalc),
    anthroheightcalc = as.numeric(anthroheightcalc)
  ) %>%
  # Exclude cases from unused assessment waves
  filter(!is.na(eventname))

```
:::

### Descriptives

::: blue
**This code creates a descriptives table**

```{r descriptives}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

descriptives_table <- df %>%
  select(eventname, sex, race.4level, hisp, anthroweightcalc) %>%
  mutate(eventname = factor(eventname, labels = c("Baseline", "Year 1","Year 2","Year 3","Year 4"))) %>%
  mutate(sex = factor(sex, labels = c("Female", "Male"))) %>%
  tbl_summary(
    by = eventname,
    missing = "no",
    label = list(sex ~ "Sex", race.4level ~ "Race", hisp ~ "Hispanic", 
                 anthroweightcalc ~ "Weight"),
    statistic = list(all_continuous() ~ "{mean} ({sd}) )", all_categorical() ~ "{p}%"),
  ) %>%
modify_header(all_stat_cols() ~ "**{level}**<br>N = {n}") %>%
  bold_labels() %>%
  italicize_levels() %>%
  modify_spanning_header(all_stat_cols() ~ "**Assessment Wave**")
theme_gtsummary_compact()

descriptives_table
```

:::
:::

## Results

::: panel-tabset
### Build Model {.tabset .tabset-fade .tabset-pills}

::: blue
The code snippet below tells R to conduct a multiple regression analysis by subtracting each participant's height at T1 from their height at T0. Relevant summary statistics are also provided.

**Compute Multiple Regression Model**
```{r build-model-resid-regression-1}
#| class.source: 'fold-hide'
#| message: FALSE
#| warning: FALSE
#| echo: TRUE

# Split data using the correct eventname values
baseline_data <- df %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  select(src_subject_id, Height_baseline = anthroheightcalc, Weight_baseline = anthroweightcalc)

followup_data <- df %>% 
  filter(eventname == "1_year_follow_up_y_arm_1") %>% 
  select(src_subject_id, Height_followup = anthroheightcalc)

# Merge baseline and follow-up data
merged_data <- baseline_data %>%
  left_join(followup_data, by = "src_subject_id")

# Check the first few rows of the merged data
#head(merged_data)

# Fit the regression model excluding NA values
model <- lm(Height_followup ~ Height_baseline + Weight_baseline, data = merged_data, na.action = na.exclude)

```

:::

### Model Summary {.tabset .tabset-fade .tabset-pills}

::: blue

A plot to show xxxxx.
```{r model-checking}
model_summary <- summary(model)
print(model_summary)

par(mfrow = c(2, 2))  # Set up a 2x2 plotting area
plot(model)  # Generate diagnostic plots

```

This regression analysis evaluates how baseline height and weight predict children's height at a 1-year follow-up. The output from our model provides several key statistics:

i. an F-statistic of `r model_summary$fstatistic[1]`;
ii. degrees of freedom of `r model_summary$df[2]`;
iii. parameter estimates for baseline height and weight;
iv. standard errors for these estimates; and
v. p-values for the significance of these estimates.

For every one unit increase in baseline height, there's an associated increase of approximately `r round(model_summary$coefficients["Height_baseline", "Estimate"], 3)` units in the follow-up height, and this effect was statistically significant with a p-value of `r model_summary$coefficients["Height_baseline", "Pr(>|t|)"]`. Similarly, for every one unit increase in baseline weight, there's an associated increase of approximately `r round(model_summary$coefficients["Weight_baseline", "Estimate"], 4)` units in the follow-up height, which is also statistically significant with a p-value of `r model_summary$coefficients["Weight_baseline", "Pr(>|t|)"]`. Overall, this model explains a substantial portion of the variance in follow-up height, with an adjusted R-squared value of `r round(model_summary$adj.r.squared, 4)`. The overall model is highly significant with a p-value less than `r model_summary$coefficients[1,4]`, indicating that the predictors, collectively, have a significant relationship with the dependent variable.

:::

### Model Plots  {.tabset .tabset-fade .tabset-pills}

::: blue
```{r scatterplot}

# Scatterplot with regression line for Weight_baseline
ggplot(merged_data, aes(x = Weight_baseline, y = Height_followup)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "black") + 
  labs(title = "Relationship between Baseline Weight and Follow-up Height",
       x = "Baseline Weight",
       y = "Follow-up Height") +
  theme_minimal()

```

The scatterplot visually depicts the relationship between children's weights at baseline (`r names(merged_data)[3]`) and their heights at Year_1 (`r names(merged_data)[4]`). Each point on the plot represents a child, with their baseline weight plotted on the x-axis and their Year_1 height on the y-axis. The clear positive linear trend, as illustrated by the black regression line, indicates that children who had higher weights at baseline generally had higher heights at Year_1. While further statistical analyses can quantify the strength and direction of this relationship, visually, the data suggests a pronounced positive association between baseline weight and follow-up height.

:::
:::

## Wrapping Up
<div class = "blue">

::: panel-tabset
### Write-up {.tabset .tabset-fade .tabset-pills}

::: blue

The regression analysis was conducted to predict children's height at the 1-Year follow-up using their baseline height (`r names(merged_data)[2]`) and weight (`r names(merged_data)[3]`). Both predictors were statistically significant. Specifically, for every one-unit increase in baseline height, the height at Year_1 increased by approximately `r round(model_summary$coefficients["Height_baseline", "Estimate"], 3)` units, while holding weight constant. Similarly, for every one-unit increase in baseline weight, the height at Year_1 increased by about `r round(model_summary$coefficients["Weight_baseline", "Estimate"], 4)` units, while holding baseline height constant. The overall model explained a substantial `r round(model_summary$adj.r.squared * 100, 2)`% of the variance in height at the 1-Year follow-up, as indicated by the adjusted R-squared value.

:::
:::