### 2024-03-06 examples/_queue/_example_CBCLADHD.qmd

ABCD worked example when the outcome variable is highly skewed. CBCL ADHD summary score was the outcome of interest. This variable is both overdispersed and zero-inflated.

overdispersment and zero-inflation can be verified using packages like the DHARMa package in R.

glmmTMB was used to model CBCL ADHD. Alternative packages that can also be used include the brms and glmmAdapative package.

examples include a fixed effect for interview age and sex (reference = female). interview age was scaled to improve model run time.

a random intercept was included for the family structure. A random intercept was not included for study site after review of the intraclass correlation coefficient.

a random slope was included for the study visit. A fixed effect was also included for study visit.

zero inflated models included the same covariates.

parallel processing was used to speed up the model run time.

comparison of different distributions is not included here.

dataset annual_ABCD includes only participants that completed all five annual visits included in this example.

```{r}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE
#| cache: FALSE

library(dplyr)

# Set the data paths
data_path_1 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/demo5.0.rds"
data_path_2 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/core-rds-5.0/non-imaging_excluding_nt_5.0.rds"

# Read the data
data_demographics <- readRDS(data_path_1)
data_nonimaging <- readRDS(data_path_2)

data_nonimaging_r <-  data_nonimaging %>%
  filter(eventname == "baseline_year_1_arm_1" | 
           eventname == "1_year_follow_up_y_arm_1" |
           eventname == "2_year_follow_up_y_arm_1" | 
           eventname == "3_year_follow_up_y_arm_1" |
           eventname == "4_year_follow_up_y_arm_1")

# Subset the nonimaging data to include desired variables
selected_vars <- c("src_subject_id", "eventname",
                   "cbcl_q04_p","cbcl_q08_p","cbcl_q10_p",
                   "cbcl_q41_p","cbcl_q78_p","cbcl_q93_p","cbcl_q104_p")

subset_data <- data_nonimaging_r[, selected_vars]

#below requires the demo5.0 dataframe to extract the sex variable
xy1 <- data_demographics[,c("src_subject_id", "eventname", "sex", "age","rel_family_id.bl")]

annual_ABCD <-  merge(subset_data, xy1, all.x =TRUE, by=c("eventname", "src_subject_id"))

#Manual creation of the CBCL ADHD variable. May need to convert these to numeric first.These
#were originally factor variables that I manually mutated using dplyr.

# Assuming annual_ABCD is your data frame
annual_ABCD$CBCL_ADHD <- rowSums(annual_ABCD[c("cbcl_q04_p", "cbcl_q08_p",
                                               "cbcl_q10_p", "cbcl_q41_p", 
                                               "cbcl_q78_p", "cbcl_q93_p", 
                                               "cbcl_q104_p")], na.rm = TRUE)

#making a numeric study visit variable
annual_ABCD<- annual_ABCD %>% 
  mutate(visit = case_when(
    eventname == "baseline_year_1_arm_1" ~ 1,
    eventname == "1_year_follow_up_y_arm_1" ~ 2,
    eventname == "2_year_follow_up_y_arm_1" ~3,
    eventname == "3_year_follow_up_y_arm_1" ~ 4,
    eventname == "4_year_follow_up_y_arm_1" ~ 5))

###removing any participants with fewer than 5 CBCL ADHD datapoints
annual_ABCD <- annual_ABCD %>% 
  group_by(src_subject_id) %>% 
  filter(n() >= 5)

```

## 
```{r}

library(glmmTMB)

#Generalized Poisson Model
model_GenPois <- glmmTMB(CBCL_ADHD ~ 1 +  age + sex + 
                           factor(visit)+(1+visit| rel_family_id.bl),
                         family=genpois(link="log"), data=annual_ABCD, 
                         REML=FALSE, control = glmmTMBControl(parallel = 16))

summary(model_GenPois)

```

```{r}

confint(model_GenPois)
```

Interpretation:In the model without the ZI term, a 1-SD increase in interview age was associated with a -0.03 (95% CI: -0.08, 0.024, p-value=0.27) decrease in CBCL ADHD score. This was non-significant. Being male (reference=female) was associated with a 0.445-point increase in CBCL ADHD score (95% CI: 0.39, 0.5, p-value <0.0001).

## Truncated Generalized Poisson Model
```{r}

system.time(Model_ZIGenPois <- glmmTMB(CBCL_ADHD ~ 1 +  age + sex +
                                         factor(visit) + (1+visit| rel_family_id.bl),
                                       family=truncated_genpois(link="log"),
                                       ziformula=~1 + age + sex + factor(visit) +
                                         (1+visit| rel_family_id.bl),
                                       data=annual_ABCD, REML=FALSE, control =
                                         glmmTMBControl(parallel = 16)))

summary(Model_ZIGenPois)
```

```{r}
confint(model_GenPois)
```

Interpretation: In the model with the ZI term: In those with a CBCL ADHD score > 0, a 1-SD increase in interview age was associated with -0.002-point decrease in CBCL ADHD score (95% -0.05, 0.044, p value = 0.94). In those with a CBCL ADHD score > 0, being male was associated with a 0.27-point increase in CBCL ADHD score (95% CI: 0.22, 0.31, p-value < 0.0001). In the model with the ZI term: In those with a CBCL ADHD score = 0, a 1-SD increase in interview age was associated with 0.17-point increase in CBCL ADHD score (95% 0.018, 0.32, p value = 0.0282). In those with a CBCL ADHD score = 0, being male was associated with a -0.86-point decrease in CBCL ADHD score (95% CI: -1.00, -0.71, p-value < 0.0001).