### 2024-03-12 examples/web-r/webr-test.qmd
updates

---
title: webR in Quarto HTML Documents
format: html
engine: knitr
# webr: 
#   show-startup-message: false    # Disable displaying status of webR initialization
#   packages: ['ggplot2', 'dplyr'] # Install R packages on document open
# 'tidyverse', 'easystats', 'lme4', 'gtsummary', 'report', 'broom', 'gridExtra')
filters:
  - webr
---

This is a webr-enabled code cell in a Quarto HTML document.

```{webr-r}
fit <- lm(mpg ~ am, data = mtcars)
summary(fit)
```

```{webr-r}

# Set the data paths
data_path_1 <- "FaKEbcd.rds"

# Read the data
df <- readRDS(data_path_1)

head(df)
str(df)

```