### 2024-05-01 Template.qmd
update unconditional examples

---
title: "Template"
author: "author name"
---

```{r}
#| echo: FALSE # The content of this code block will not be shown when rendered
#| output: FALSE # Output from this code block will not be shown when rendered

# This code configures knitr code chunk options

knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, output=T,
                      comment=NA, cache=T, code_folding=T,
                      R.options=list(width=220, digits = 3),
                      fig.align='center', 
                      out.width='75%', fig.asp=.75)
```

```{r}
#| echo: FALSE # The content of this code block will not be shown when rendered
#| output: FALSE # Output from this code block will not be shown when rendered
#| cache: FALSE

# This code loads the data to be used for the example

[Enter code to load example data]

```

## Overview
[Introductory text for the example]

## Preliminary Setup

### Load R libraries
```{r}

library(xxxxx)    # xxxxxxx
library(xxxxx)    # xxxxxxx

```

### Descriptives

```{r}
[Code for providing sample descriptives]
```

## Results

### [Name of First Set of Analyses]
[Concise description of the analysis approach]

```{r}

[Enter analysis code]

```

[Results write-up for analysis]

### Model Plots
[Concise description of the plot type]

```{r}
[Enter plot code]
```
[Concise summary of plot output]

## Wrapping Up

[Concise summary of the example]