### 2024-05-01 examples/3b_minimalist_Examples_LinearMixedModels.qmd

---
title: "Linear Mixed Models: Random Intercept and Slope"
author: "Biostatistics Working Group"
---

```{r}
#| echo: FALSE # The content of this code block will not be shown when rendered
#| output: FALSE # Output from this code block will not be shown when rendered

# This code configures knitr code chunk options

knitr::opts_chunk$set(
    echo = T, message = F, warning = F, error = F, output = T,
    comment = NA, cache = T, code_folding = T,
    R.options = list(width = 220, digits = 3),
    fig.align = "center",
    out.width = "75%", fig.asp = .75
)
```

```{r}
#| echo: FALSE # The content of this code block will not be shown when rendered
#| output: FALSE # Output from this code block will not be shown when rendered
#| cache: FALSE

# This code loads the data to be used for the example

library(tidyverse)

# Set the data path
data_path <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/core-rds-5.0/non-imaging_excluding_nt_5.0.rds"

# Read the data
df <- readRDS(data_path)

# Create and mutate the dataframe
df_long <- df %>%
    select(src_subject_id, eventname, cbcl_scr_syn_external_r) %>%
    filter(eventname %in% c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1")) %>%
    drop_na(cbcl_scr_syn_external_r) %>%
    arrange(src_subject_id, eventname) %>%
    mutate(
        src_subject_id = factor(src_subject_id),
        eventname = factor(eventname, levels = c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1"), labels = c("Baseline", "Year_1", "Year_2", "Year_3")),
        cbcl_scr_syn_external_r = round(as.numeric(cbcl_scr_syn_external_r), 2)
    ) %>%
    rename(
        id = src_subject_id,
        event = eventname,
        height = cbcl_scr_syn_external_r
    )

df_long <- df_long %>%
  group_by(id) %>%
  filter(all(c("Baseline", "Year_1", "Year_2", "Year_3") %in% event)) %>%
  ungroup()

```

## Overview

The Linear mixed model (LMM) with a random intercept and slope extends the traditional fixed-effect linear regression by incorporating random effects for the intercept and slope parameters. This allows each participant to have their own unique intercept and slope values, reflecting individual starting points and rates of change, in addition to the overall mean-level (fixed-effect) trajectory.

This example will examine trajectories of height obtained across 3 annual measurement occasions in a sample of youth taking part in the ABCD Study. The primary aim is to characterize stability and change in participant's height, while accounting for observations that are clustered within youth over time. The LMM with a random intercept and slope will simultaneously estimate an overall sample mean (fixed effect) for the height trajectory's intercept and slope values, as well as subject-specific (random) effects that vary randomly about the mean intercept and slope values.

## Preliminary Setup
### Load Packages
```{r}

library(tidyverse) # Collection of R packages for data science
library(gtsummary) # Publication-ready tables
library(rstatix) # Statistical Tests in tidy format
library(lme4) # Linear mixed-effects models
```

### Descriptives
```{r}
#| echo: FALSE # The content of this code block will not be shown when rendered
#| output: FALSE # Output from this code block will not be shown when rendered
#| cache: FALSE

# This code loads the data to be used for the example

library(tidyverse)

# Set the data path
data_path <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/core-rds-5.0/non-imaging_excluding_nt_5.0.rds"

# Read the data
df <- readRDS(data_path)

# Create and mutate the dataframe
df_long <- df %>%
    select(src_subject_id, eventname, cbcl_scr_syn_external_r) %>%
    filter(eventname %in% c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1")) %>%
    drop_na(cbcl_scr_syn_external_r) %>%
    arrange(src_subject_id, eventname) %>%
    mutate(
        src_subject_id = factor(src_subject_id),
        eventname = factor(eventname, levels = c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1"), labels = c("Baseline", "Year_1", "Year_2", "Year_3")),
        cbcl_scr_syn_external_r = round(as.numeric(cbcl_scr_syn_external_r), 2)
    ) %>%
    rename(
        id = src_subject_id,
        event = eventname,
        height = cbcl_scr_syn_external_r
    )

df_long <- df_long %>%
  group_by(id) %>%
  filter(all(c("Baseline", "Year_1", "Year_2", "Year_3") %in% event)) %>%
  ungroup()

```

## Results 
### Compute LMM with Random Intercepts and Slopes
This code fits a random-intercept and random-slope LMM to predict 'Height' across 3 annual visits, accounting for individual-level variability (random effects) of intercept and slope values across participants. Model output is provided below, as well as a brief summary of results.

This model accounts for individual variability in height by including a random intercept and sllope for each subject (id). The output indicates that the model was fit using the REML (Restricted Maximum Likelihood) criterion.

```{r}

model <- lmer(height ~ event + (1 + event | id), data = df_long, REML = T)

# Model with random intercepts only
model_ri_only <- lmer(height ~ event + (1 | id), data = df_long, REML = T)
model_ri_only

# Model with random slopes only, without intercepts
model_rs_only <- lmer(height ~ event + (0 + event | id), data = df_long, REML = T)
model_rs_only

## Output and summary statistics
summary(model)
confint(model, level = 0.95, method = "Wald")

```

These results show an increasing trajectory of height across the years, with an increase of approximately 2.36 inches in Year 1 and 5.02 inches by Year 2 compared to the baseline, as evidenced by the fixed effects estimates. The random intercept and slope effects suggests some between-subject variability in particpant's height values at baseline and in rates of change over time. The model's residuals are mostly well-behaved, although the range suggests a few extreme values or outliers.

### Model Plots
The following set of plots are used to facilitate model diagnostics. The first is a histogram showcasing the distribution of random intercepts for individual subjects, indicating variations in height not explained by the fixed effects. The second depicts residuals versus fitted values, helping assess the model's fit and potential heteroscedasticity. The third contrasts observed and predicted height values across different time points, offering a side-by-side evaluation of the model's predictions against actual observations.

```{r}

# 1. Extract the random effects
random_effects <- ranef(model)[[1]]

# 2. Convert to dataframe
random_effects_df <- data.frame(Intercept = random_effects$`(Intercept)`)

# 3. hist_plot
hist_plot <- ggplot(random_effects_df, aes(x = Intercept)) +
    geom_histogram(aes(y = after_stat(density)), bins = 30, color = "black", fill = "lightblue") +
    labs(title = "Histogram of Random Effects", x = "Random Intercept Values", y = "Density") +
    theme_minimal()

print(hist_plot)

```

```{r}

# Extract fixed effects
fixed_effects <- fixef(model)

# Prepare data frame for plotting
df_long$predicted_height <- predict(model, re.form = NA) # Overall fixed effect predictions

# Calculate individual intercept adjustments
df_long$individual_intercept <- predict(model, re.form = ~ (1 | id)) - fixed_effects[1]

# Generate the plot
ggplot(df_long, aes(x = event, y = height, group = id)) +
    geom_line(aes(y = height, color = "Individual Trajectories"), alpha = 0.3) +
    geom_line(data = df_long, aes(x = event, y = predicted_height, group = 1, color = "Mean Trajectory"), linewidth = 1.5) +
    scale_color_manual(values = c("Individual Trajectories" = "grey", "Mean Trajectory" = "blue")) +
    labs(
        title = "Individual and Group-Mean Height Trajectories",
        x = "Event Name",
        y = "Height (cm)",
        color = "Trajectory Type"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom")

```

This plot illustrates the change in participant's height across 3 annual measurement occassions. The individual raw data points are shown as red lines, representing the original height measurements, which demonstrate the variability among participants at each time point. These are overlaid with grey lines, which adjust each individual's trajectory unique growth pattern indicated by the fixed effects from the model. The blue line represents the sample mean trajectory, calculated from the model's fixed effects, showcasing the average growth trend across all participants.

The plot shows individual trajectories that vary around the model's predicted values due to random effects, indicating substantial inter-individual variability in baseline heights and rates of change in height over tim. On average, participants experienced an increase in height of approximately 2.36 inches from baseline to Year 1 and an additional 2.66 inches from Year 1 to Year 2. The random effects for the estimated intercept and slope suggest differences in starting heights and rates of change over time, reflecting the variability in growth patterns among the participants.

## Wrapping Up

The random-intercept and random-slope LMM analysis revealed an increasing pattern of change in children's height measured over 3 annual assessments. The mean intercept value showed an average initial height of 55.24 inches (sd = 3.03) with evidence of individual variability around this (random effect) estimate. An examination of the slope factor revealed year-to-year growth increments of approxmiately 2.5 inches, with evidence of individual variability around this estimate.

The LMM framework used in this example offered the ability to account for individual differences in initial height and changes in growth over time, demonstrating its utility in handling clustered or correlated data that arise from repeated measurements on the same subjects. More generally, this approach is often particularly useful in developmental studies as it allows for the inclusion of both fixed effects, which estimate common parameters across all individuals, and random effects, which capture unique variations among individuals. This enables a more precise understanding of both the group-level trends and individual-specific trajectories in longitudinal data.