### 2024-05-01 examples/_queue/_13b_Examples_LatentClassGrowthAnalsis.qmd

---
title: "Latent Class Growth Analysis"
author: "Biostatistics Working Group"
---

# LCGA: {.unnumbered}

## Overview

<p>

"A valuable tool for investigating heterogeneity in symptom patterns is Latent Growth Curve Modelling (GCM), a statistical method that analyses inter-individual variability in intra-individual patterns over time @curran2010. One specific type of GCM is Latent Class Growth Analysis (LCGA), which considers unobserved heterogeneity (different groups) over time within a larger population @jung2008, @nguena2020, @ram2009. In other words, LCGA can examine the growth and shape of the course of depressive symptoms over time and assess how individuals in the population group together based on their symptom patterns."

"The aim of the present study is to investigate the course of depressive symptoms over a period of xx years .... and to .... . To achieve this goal, LGCM will be performed on data from the xxxx ABCD cohort (add citation). The ABCD study was initiated in xxxxx. Based on previous literature @musliner2016, we hypothesize that three or four trajectories of depression can be identified. Given that xxxxx, we additionally hypothesize that individuals following more xxxx trajectories will be more/less likely to experience xxxx and exhibit reduced levels of xxxxx"

</p>

## Preliminary Setup

::: panel-tabset
### Install Packages {.tabset .tabset-fade .tabset-pills}

::: blue
> **This code installs the r packages necessary for this example, if
> they are not already installed**

```{r pckg-install}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

# Create a list of required packages
packages_required <- c("tidyverse","tidySEM","rstatix","report","broom","gridExtra","easystats","gtsummary")

# Check which packages are not installed and install them
packages_to_install <- setdiff(packages_required, rownames(installed.packages()))
if (length(packages_to_install) > 0) {
    install.packages(packages_to_install)
}

# Load the required packages
lapply(packages_required, library, character.only = TRUE)

```
:::

### Load Packages

::: blue
**This code loads the r libraries necessary for this example**

```{r lib-load}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

library(tidyverse)    # Collection of R packages for data science
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(tidySEM)      # Structural equation modeling in R
library(report)       # Easy reporting of regression analyses
library(broom)        # Tidy and augment statistical models output
library(gridExtra)    # Arrange multiple grid-based plots on a page
library(easystats)    # Descriptive statistics and robust statistical methods
library(gtsummary)    # Presentation-ready data summary and analytic result tables
```
:::

### Config Options

::: blue
**This code configures knitr code chunk options**

```{r config}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, 
                      comment=NA, cache=T, code_folding=T,
                      R.options=list(width=220, digits = 3),
                      fig.align='center', 
                      out.width='75%', fig.asp=.75)
```
:::
:::

## Descriptives Overview

::: panel-tabset
### Read and View Data {.tabset .tabset-fade .tabset-pills}

::: blue
**This code reads in and shows the data to be used in the current
example**

```{r read-data}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE
#| cache: FALSE

# Set the data paths
data_path_1 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/demo5.0.rds"
data_path_2 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/core-rds-5.0/non-imaging_excluding_nt_5.0.rds"

# Read the data
data_demographics <- readRDS(data_path_1)
data_nonimaging <- readRDS(data_path_2)

# Subset the nonimaging data to include desired variables
selected_vars <- c("src_subject_id", "eventname", "nihtbx_totalcomp_fc", "anthroweightcalc", "anthroheightcalc")
subset_data <- data_nonimaging[, selected_vars]

library(dplyr)
# # Merge the datasets on 'src_subject_id' and 'eventname'
merged_data <- data_demographics %>%
  full_join(subset_data, by = c("src_subject_id", "eventname"))

# Inspect the merged data structure
str(merged_data)

# Define event names to be retained in the analysis and convert variables to appropriate data types
eventnames_to_include <- c("baseline_year_1_arm_1",
                           "1_year_follow_up_y_arm_1",
                           "2_year_follow_up_y_arm_1",
                           "3_year_follow_up_y_arm_1",
                           "4_year_follow_up_y_arm_1")

df <- merged_data %>%
  filter(eventname %in% eventnames_to_include) %>%
  mutate(
    src_subject_id = as.factor(src_subject_id),
    eventname = factor(eventname, levels = eventnames_to_include, ordered = TRUE),
    age = as.numeric(age),
    sex = as.factor(sex),
    race.4level = as.factor(race.4level),
    hisp = as.factor(hisp),
    high.educ.bl = as.factor(high.educ.bl),
    household.income.bl = as.factor(household.income.bl),
    acs_raked_propensity_score = as.numeric(acs_raked_propensity_score),
    rel_family_id.bl = as.factor(rel_family_id.bl),
    site_id_l = as.factor(site_id_l),
    nihtbx_totalcomp_fc = as.numeric(nihtbx_totalcomp_fc),
    anthroweightcalc = as.numeric(anthroweightcalc),
    anthroheightcalc = as.numeric(anthroheightcalc)
  ) #%>%
#   # Exclude cases from unused assessment waves
#   filter(!is.na(eventname))

```
:::

```{r}

# Long to wide conversion
df_long <- df %>%
  select(src_subject_id, eventname, anthroheightcalc) %>% na.omit()

# Creating df_wide
df_wide <- pivot_wider(
    data = df_long,
    names_from = eventname,
    values_from = anthroheightcalc,
    id_cols = src_subject_id
)

df_wide <- df_wide %>%
    rename(
        anthroheight.0 = baseline_year_1_arm_1,
        anthroheight.1 = `1_year_follow_up_y_arm_1`,
        anthroheight.2 = `2_year_follow_up_y_arm_1`,
        anthroheight.3 = `3_year_follow_up_y_arm_1`,
        anthroheight.4 = `4_year_follow_up_y_arm_1`
    )

## Subset the nonimaging data to include desired variables
selected_vars_wide <- c("anthroheight.0", "anthroheight.1", "anthroheight.2", "anthroheight.3", "anthroheight.4")
df_wide <- df_wide[, selected_vars_wide]
str(df_wide)

```

### Descriptives

::: blue
**This code creates a descriptives table**

```{r descriptives}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

descriptives_table <- df %>%
  select(eventname, sex, race.4level, hisp, anthroweightcalc) %>%
  mutate(eventname = factor(eventname, labels = c("Baseline", "Year 1","Year 2","Year 3","Year 4"))) %>%
  mutate(sex = factor(sex, labels = c("Female", "Male"))) %>%
  tbl_summary(
    by = eventname,
    missing = "no",
    label = list(sex ~ "Sex", race.4level ~ "Race", hisp ~ "Hispanic", 
                 anthroweightcalc ~ "Weight"),
    statistic = list(all_continuous() ~ "{mean} ({sd}) )", all_categorical() ~ "{p}%"),
  ) %>%
modify_header(all_stat_cols() ~ "**{level}**<br>N = {n}") %>%
  bold_labels() %>%
  italicize_levels() %>%
  modify_spanning_header(all_stat_cols() ~ "**Assessment Wave**")
theme_gtsummary_compact()

descriptives_table
```
:::

:::

We can examine these distributions visually as well:

```{r}
#| echo: FALSE
#| messages: FALSE
#| warning: FALSE
#| output: TRUE

# df_long Plots
ggplot(df_long, aes(x = anthroheightcalc)) +
    geom_density() +
    facet_wrap(~eventname) +
    theme_bw()

```

As this type of skew can result in convergence problems in LCGA, we compared several transformations to reduce skew: The square and cube root, log, inverse, and Box-Cox transformations.

```{r}
#| echo: FALSE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

df_scores <- df_long
# Store original range of cbcl_scr_dsm5_depress_t
rng_height <- range(df_scores$anthroheightcalc, na.rm = TRUE)
# Log-transform
df_scores$log <- scales::rescale(log(df_scores$anthroheightcalc), to = c(
    0,
    1
))
# Square root transform
df_scores$sqrt <- scales::rescale(sqrt(df_scores$anthroheightcalc), to = c(
    0,
    1
))
# Cube root transform
df_scores$qrt <- scales::rescale(df_scores$anthroheightcalc^0.33, to = c(
    0,
    1
))
# Reciprocal transform
df_scores$reciprocal <- scales::rescale(1 / df_scores$anthroheightcalc, to = c(
    0,
    1
))

# # Define function for Box-Cox transformation
# bc <- function(x, lambda) {
#     (((x^lambda) - 1) / lambda)
# }

# # Inverse Box-Cox transformation
# invbc <- function(x, lambda) {
#     ((x * lambda) + 1)^(1 / lambda)
# }
# # Box-Cox transform
# b <- MASS::boxcox(lm(df_scores$anthroheightcalc ~ 1), plotit = FALSE)
# lambda <- b$x[which.max(b$y)]
# df_scores$boxcox <- bc(df_scores$anthroheightcalc, lambda)
# # Store range of Box-Cox transformed data
# rng_bc <- range(df_scores$boxcox)
# df_scores$boxcox <- scales::rescale(df_scores$boxcox, to = c(
#     0,
#     1
# ))

# Rescale anthroheightcalc
df_scores$anthroheightcalc <- scales::rescale(df_scores$anthroheightcalc, to = c(0, 1))

```

We can plot these transformations:

```{r}
#| echo: FALSE
#| messages: FALSE
#| warning: FALSE
#| output: TRUE

# Make plot data
df_plot <- do.call(rbind, lapply(c(
    "anthroheightcalc", "log", "sqrt", "qrt" #,
    #"boxcox"
), function(n) {
    data.frame(df_scores[c("eventname", "src_subject_id")],
        Value = df_scores[[n]],
        Transformation = n
    )
}))
# Plot
ggplot(df_plot, aes(x = Value, colour = Transformation)) +
    geom_density() +
    facet_wrap(~eventname) +
    scale_y_sqrt() +
    xlab("Height (rescaled to 0-1)") +
    theme_bw()

```

The Box-Cox transformation (maybe/slightly?) reduced skew the most. Consequently, we proceeded with the Box-Cox transformed scores for analysis.

```{r}
#| echo: FALSE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

dat <- df_scores[, c("src_subject_id", "eventname", "log")]
dat <- reshape(dat,
    direction = "wide", v.names = "log", timevar = "eventname",
    idvar = "src_subject_id"
)

dat <- na.omit(dat)

names(dat) <- gsub("log.", "anthroheightcalc", names(dat))
```

:::

## Results

::: panel-tabset
### Build Model {.tabset .tabset-fade .tabset-pills}

::: blue
The code fits an LCGA to examine the 'Height' variable across time points ('eventname'). It xxxxxx. The results of the model are then printed to provide a detailed summary of model parameters.

**STEP 1: Compute LCGA**
```{r}
#| echo: FALSE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

#set.seed(27796)
dat[["src_subject_id"]] <- NULL
res_step <- mx_growth_mixture(
    model = "
  i =~ 1*anthroheightcalcbaseline_year_1_arm_1 + 1*anthroheightcalc1_year_follow_up_y_arm_1 + 1*anthroheightcalc2_year_follow_up_y_arm_1 +1*anthroheightcalc3_year_follow_up_y_arm_1 +1*anthroheightcalc4_year_follow_up_y_arm_1
  step =~ 0*anthroheightcalcbaseline_year_1_arm_1 + 1*anthroheightcalc1_year_follow_up_y_arm_1 + 1*anthroheightcalc2_year_follow_up_y_arm_1 +1*anthroheightcalc3_year_follow_up_y_arm_1 +1*anthroheightcalc4_year_follow_up_y_arm_1
  s =~ 0*anthroheightcalcbaseline_year_1_arm_1 + 0*anthroheightcalc1_year_follow_up_y_arm_1 + 1*anthroheightcalc2_year_follow_up_y_arm_1 +2*anthroheightcalc3_year_follow_up_y_arm_1 +3*anthroheightcalc4_year_follow_up_y_arm_1
  anthroheightcalcbaseline_year_1_arm_1 ~~ vanthroheightcalcbaseline_year_1_arm_1*anthroheightcalcbaseline_year_1_arm_1
  anthroheightcalc1_year_follow_up_y_arm_1 ~~ vanthroheightcalc1_year_follow_up_y_arm_1*anthroheightcalc1_year_follow_up_y_arm_1
  anthroheightcalc2_year_follow_up_y_arm_1 ~~ vanthroheightcalc2_year_follow_up_y_arm_1*anthroheightcalc2_year_follow_up_y_arm_1
  anthroheightcalc3_year_follow_up_y_arm_1 ~~ vanthroheightcalc3_year_follow_up_y_arm_1*anthroheightcalc3_year_follow_up_y_arm_1
  anthroheightcalc4_year_follow_up_y_arm_1 ~~ vanthroheightcalc4_year_follow_up_y_arm_1*anthroheightcalc4_year_follow_up_y_arm_1
  i ~~ 0*i
  step ~~ 0*step
  s ~~ 0*s
  i ~~ 0*s
  i ~~ 0*step
  s ~~ 0*step",
    classes = 1:3, data = dat
)
# Additional iterations because of convergence problems for
# model 1:
res_step[[1]] <- mxTryHardWideSearch(res_step[[1]], extraTries = 50)
```

```{r}
#| echo: FALSE
#| messages: FALSE
#| warning: FALSE
#| output: TRUE

# Get fit table fit
tab_fit <- table_fit(res_step)
# Select columns
tab_fit[, c("Name", "Classes", "LL", "Parameters", "AIC", "BIC", "saBIC", "Entropy", "prob_min", "n_min")]

# , "warning", "lmr_p")]

```

```{r}
#| echo: FALSE
#| messages: FALSE
#| warning: FALSE
#| output: TRUE

plot(tab_fit, statistics = c("AIC", "BIC", "saBIC"))
```

```{r}
#| echo: FALSE
#| messages: FALSE
#| warning: FALSE
#| output: TRUE

res_final <- mx_switch_labels(res_step[[3]],
    param = "M[1,7]",
    decreasing = FALSE
)
tab_res <- table_results(res_final, columns = NULL)
# Select rows and columns
tab_res <- tab_res[
    tab_res$Category %in% c("Means", "Variances"),
    c("Category", "lhs", "est", "se", "pval", "confint", "name")
]
tab_res
```

```{r}
#| echo: FALSE
#| messages: FALSE
#| warning: FALSE
#| output: TRUE

names(coef(res_final))
```
```{r}
#| echo: FALSE
#| messages: FALSE
#| warning: FALSE
#| output: TRUE

wald_tests <- wald_test(res_final, "
                   class1.M[1,6] = class2.M[1,6] &
                   class1.M[1,6] = class3.M[1,6] ;
                   class1.M[1,7] = class2.M[1,7] &
                   class1.M[1,7] = class3.M[1,7] ;
                   class1.M[1,8] = class2.M[1,8] &
                   class1.M[1,8] = class3.M[1,8]")

# Rename the hypothesis
wald_tests$Hypothesis <- c("Mean i", "Mean step", "Mean slope")
knitr::kable(wald_tests, digits = 2, caption = "Wald tests")

```

```{r}
# Assuming rng_height and lambda are defined previously in your script
brks <- seq(0, 1, length.out = 5) # Breaks in the rescaled 0-1 range

# Convert breaks back to the original log scale
# Note: This assumes that the original transformation was simply log(x) scaled to 0-1
labs <- exp(scales::rescale(brks, from = c(0, 1), to = rng_height))

p <- plot_growth(res_step[[3]], rawdata = TRUE, alpha_range = c(0, 0.05))
p <- p + scale_y_continuous(
    breaks = brks, # Use the same breaks for simplicity
    labels = round(labs, 2) # Round the labels for readability
) + ylab("Height (rescaled from log)")
p

```

#### Covariates

Assessment of covariates using the three-step approach revealed that the trajectories did not differ in any of the demogra phical variables, including sex (ΔLL(5) = 6.3 , p = .27, pcorr = .37), age (ΔLL(8) = 8.4 , p = .40, pcorr = .50), rank Journal of Affective Disorders 354 (2024) 702--711 707(ΔLL(11) = 13.5 , p = .26, pcorr = .37), educational level (ΔLL(9) = 7.1 , p = .63, pcorr = .66), function (ΔLL(8) = 8.1 , p = .43, pcorr = .51). Significant disparities in xxxx were observed, with the xxxxx group exhibiting lower counts of xxxxx compared to the xxxxx group (ΔLL(2) = 23.0 , p \< .01, pcorr \< .01), xxxxx (ΔLL(2) = 325.7 , p \< .01, pcorr \< .01), and xxxxx group (ΔLL(2) = 9.5 , p \< .01, pcorr \< .05) (Supplementary Fig. Sxx). A detailed breakdown of percentages per xxxxx is provided in Supplementary Table Sxx

#### Depression and GDT

#### Depression and substance use

## Add additional analyses section

Note that the observed individual trajectories show very high variability within classes.

:::

### Model Plots

::: blue
The following set of plots are xxxxxxx.

:::

## Wrapping Up
<div class = "blue">

::: panel-tabset
### Write-up {.tabset .tabset-fade .tabset-pills}

::: blue

The lcga was conducted to xxxxxx.

:::
:::

---------------

# Discussion

```{r}
#| echo: FALSE
#| messages: FALSE
#| warning: FALSE
#| output: TRUE

lorem::ipsum(5, sentences = c(3))
# lorem::ipsum(2, avg_words_per_sentence = 4)

```